On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-05 arity1.tex
made the section a bit more concise

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{./Fig/CI_Numbers}
	\caption{Quadratic length transcript folding deterministically into pyramid shape. Seed: thick black path. Transcript: thin blue path. Bonds: dashed red lines. }
	\label{CI:big}
\end{figure}

@@

\begin{enumerate}
	\item bead $i+1$ is fixed at $(2,1)$ and can bond with a bead at $(2,0)$ (see Fig.~\ref{fig:hexagonOut2}). Now consider bead $i+2$. For $i+1$ to be fixed at $(2,1)$, $i+2$ needs to form a bond somewhere, otherwise $i+2$ could go to $(2,1)$ forming the bond with the bead at $(2,0)$ and there would be two conformations with the maximal $1$ bond. The only possibility is that there is a bead at $(3,0)$ and $i+2$ can bond with it when placed at $(3,1)$. We can apply the same argument inductively: there is some $m\geq 0$ such that grid points $(\ell,0)$ are occupied by beads with free hands, for all $\ell\in \{2,\dots,2+m\}$, and there is no bead at $(3+m,0)$. Such an $m$ exists, and it is not greater than $n$, because those beads are all stabilized along the same side of $\hexagon_p^n$. Then, bead $i+\ell$ is fixed at $(\ell+1,1)$ and bonds with $(\ell+1,0)$. However, bead $i+2+m$ cannot be fixed anywhere, because $i+2+m$ and $i+3+m$ can only add one bond to the conformation, and that is possible either with $i+2+m \rightarrow (2+m,1)$, $i+3+m \rightarrow (3+m,1)$ or with $i+2+m \rightarrow (2+m,2)$, $i+3+m \rightarrow (2+m,1)$. Intuitively, when we reach a corner of the hexagon $\hexagon_p^n$, the next bead of the transcript cannot deterministically stabilize, as depicted in Fig.~\ref{fig:hexagonOut2}. In this case, the size of the transcript which was deterministically stabilized is bounded by the size of $\hexagon_p^n$ plus the length of one side of $\hexagon_p^{n+1}$, so by $(3n^2+3n+1)+(n+1)=3n^2+4n+2$.
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.9\linewidth]{./Fig/hexagonOut2}
		\caption{When bead $i+2+m$ is fixed}
		\label{fig:hexagonOut2}
	\end{figure}

	\item bead $i+1$ is fixed at $(0,1)$. This is only possible if
	\begin{enumerate}
		\item there is an inactive bead at $(-1,0)$ and one with a free hand one at $(-2,0)$. This case is symmetrical to (1).
		\item there is no bead at $(-1,0)$, bead $i+1$ can bond with bead $i-1$ at $(0,0)$ and the bead $i+2$ can be placed at $(-1,0)$ where it can bond with $(-2,0)$, $(-2,-1)$ or $(-1,-1)$. This leads to nondeterminism, because placing bead $i$ at $(-1,0)$ and bead $i+1$ at $(0,1)$ would yield two bonds, just as the original conformation.
		\item there is a bead at $(-1,0)$ and bead $i+1$ can bond with that or with bead $i-1$ at $(0,0)$. However, this means that placing bead $i$ at $(0,1)$ at bead $i+1$ at $(1,1)$ creates the same number of hydrogen bonds, thus resulting in bead $i$ not being placed deterministically.

	\end{enumerate}
\end{enumerate}

@@

As before, let us assume that the last bead stabilized within $\hexagon_p^n$ is at point $(0,0)$ and the next bead is the first to be stabilized outside $\hexagon_p^n$ at point $(1,1)$. Depending on whether bead $i-1$ at point (0,0) has a free hand to form a bond or not, we distinguish two cases.
\begin{enumerate}
\item  Bead $i-1$ at (0,0) has a free hand.\\ If the most stable conformation formed by beads $i, i+1, i+2$ adds only two  bonds, it will be nondeterministic because there are at least two possibilities, see Fig.~\ref{fig:beadi}. Therefore, it needs to make three new bonds to deterministically stabilize $i$.\\

@@

\item Bead $i-1$ at (0,0) does not have a free hand.\\
(i) First, let us assume that there is a bead at (1,0) which has a free hand. If there is a bead at (1,0) and beads $i, i+1, i+2$ add only two bonds, we instantly get nondeterministism as in Fig.~\ref{fig:beadat(1,0)}. Hence, beads $i, i+1, i+2$ need to make three bonds to stabilize $i$, but if they make three hydrogen bonds, the situation is analogous to (a) and (c) in Fig.~\ref{fig:3bonds}.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
        \fill (0.0, 0.0) circle [radius = 0.1];
        \fill (1.0, 0.0) circle [radius = 0.1];
        \fill (0.5, 0.866) circle [radius = 0.1];
        \fill (1.5, 0.866) circle [radius = 0.1];
        \fill (1.0, 1.732) circle [radius = 0.1];
        \fill (0.0, 0.0) node [below] {$(0, 0)$};
        \fill (0.5, 0.866) node [left] {$(1, 1)$};
        \fill (1.0, 0.0) node [below] {$(1, 0)$};
        \draw (0.0, 0.0) -- (0.5, 0.866);
        \draw (0.5, 0.866) -- (1.5, 0.866);
        \draw (1.0, 1.732) -- (1.5, 0.866);
        \draw [dashed] [red] (1.5, 0.866) -- (1.0, 0.0);
        \draw [dashed] [red] (1.0, 1.732) -- (0.5, 0.866);
        \draw [dashed] [red] (1.0, 1.732) -- (0.5, 0.866);
      \end{tikzpicture}

@@

    \end{tabular}
    \caption{Two conformations when a bead at (1,0) has a free hand}
    \label{fig:beadat(1,0)}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
        \fill (0.0, 0.0) circle [radius = 0.1];
        \fill (0.5, 0.866) circle [radius = 0.1];
        \fill (-0.5, 0.866) circle [radius = 0.1];
        \fill (0.0, 1.732) circle [radius = 0.1];
        \fill (0.0, 0.0) node [below] {$(1, 1)$};
        \draw (0.0, 0.0) -- (0.5, 0.866);
        \draw (0.5, 0.866) -- (-0.5, 0.866);
        \draw (0.0, 1.732) -- (-0.5, 0.866);
        \draw [dashed] [red] (0.0, 1.732) -- (0.5, 0.866);
        \draw [dashed] [red] (0.0, 0.0) -- (-0.5, 0.866);
      \end{tikzpicture}

@@

    \end{tabular}
    \caption{Two conformations when a bead $i$ at (1,1) has a free hand}
    \label{fig:beadi2}
  \end{center}
\end{figure}

@@

\end{enumerate}

>>>>

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{./Fig/CI_Numbers}
	\caption{Quadratic length transcript folding deterministically into pyramid shape. Seed: thick black path. Transcript: thin blue path. Bonds: dashed red lines. }
	\label{CI:big}
\end{figure}

@@

%\begin{enumerate}
\noindent\textbf{Case 1.} bead $i+1$ is fixed at $(2,1)$ and can bond with a bead at $(2,0)$ (see Fig.~\ref{fig:hexagonOut2}). Now consider bead $i+2$. For $i+1$ to be fixed at $(2,1)$, $i+2$ needs to form a bond somewhere, otherwise $i+2$ could go to $(2,1)$ forming the bond with the bead at $(2,0)$ and there would be two conformations with the maximal $1$ bond. The only possibility is that there is a bead at $(3,0)$ and $i+2$ can bond with it when placed at $(3,1)$. We can apply the same argument inductively: there is some $m\geq 0$ such that grid points $(\ell,0)$ are occupied by beads with free hands, for all $\ell\in \{2,\dots,2+m\}$, and there is no bead at $(3+m,0)$. Such an $m$ exists, and it is not greater than $n$, because those beads are all stabilized along the same side of $\hexagon_p^n$. Then, bead $i+\ell$ is fixed at $(\ell+1,1)$ and bonds with $(\ell+1,0)$. However, bead $i+2+m$ cannot be fixed anywhere, because $i+2+m$ and $i+3+m$ can only add one bond to the conformation, and that is possible either with $i+2+m \rightarrow (2+m,1)$, $i+3+m \rightarrow (3+m,1)$ or with $i+2+m \rightarrow (2+m,2)$, $i+3+m \rightarrow (2+m,1)$. Intuitively, when we reach a corner of the hexagon $\hexagon_p^n$, the next bead of the transcript cannot deterministically stabilize, as depicted in Fig.~\ref{fig:hexagonOut2}. In this case, the size of the transcript which was deterministically stabilized is bounded by the size of $\hexagon_p^n$ plus the length of one side of $\hexagon_p^{n+1}$, so by $(3n^2+3n+1)+(n+1)=3n^2+4n+2$.
	\begin{figure}[h]
		\centering
		\includegraphics[width=0.9\linewidth]{./Fig/hexagonOut2}
		\caption{When bead $i+2+m$ is fixed}
		\label{fig:hexagonOut2}
	\end{figure}

\vspace{0.2cm}

\noindent\textbf{Case 2.} bead $i+1$ is fixed at $(0,1)$. This is only possible if
	\begin{enumerate}[label=(\alph*)]
		\item there is an inactive bead at $(-1,0)$ and one with a free hand one at $(-2,0)$. This case is symmetrical to (1).
		\item there is no bead at $(-1,0)$, bead $i+1$ can bond with bead $i-1$ at $(0,0)$ and the bead $i+2$ can be placed at $(-1,0)$ where it can bond with $(-2,0)$, $(-2,-1)$ or $(-1,-1)$. This leads to nondeterminism, because placing bead $i$ at $(-1,0)$ and bead $i+1$ at $(0,1)$ would yield two bonds, just as the original conformation.
		\item there is a bead at $(-1,0)$ and bead $i+1$ can bond with that or with bead $i-1$ at $(0,0)$. However, this means that placing bead $i$ at $(0,1)$ at bead $i+1$ at $(1,1)$ creates the same number of hydrogen bonds, thus resulting in bead $i$ not being placed deterministically.

	\end{enumerate}
%\end{enumerate}

@@

As before, let us assume that the last bead stabilized within $\hexagon_p^n$ is at point $(0,0)$ and the next bead is the first to be stabilized outside $\hexagon_p^n$ at point $(1,1)$. Depending on whether bead $i-1$ at point (0,0) has a free hand to form a bond or not, we distinguish two cases.
%\begin{enumerate}
\noindent\textbf{Case 1.}  Bead $i-1$ at (0,0) has a free hand.\\ If the most stable conformation formed by beads $i, i+1, i+2$ adds only two  bonds, it will be nondeterministic because there are at least two possibilities, see Fig.~\ref{fig:beadi}. Therefore, it needs to make three new bonds to deterministically stabilize $i$.\\

@@

\noindent\textbf{Case 2.} Bead $i-1$ at (0,0) does not have a free hand.\\
(i) First, let us assume that there is a bead at (1,0) which has a free hand. If there is a bead at (1,0) and beads $i, i+1, i+2$ add only two bonds, we instantly get nondeterministism as in Fig.~\ref{fig:beadat(1,0)}. Hence, beads $i, i+1, i+2$ need to make three bonds to stabilize $i$, but if they make three hydrogen bonds, the situation is analogous to (a) and (c) in Fig.~\ref{fig:3bonds}.

@@

\begin{figure}
	\begin{minipage}{.45\textwidth}

  \begin{center}
    \begin{tabular}{cc}
      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
        \fill (0.0, 0.0) circle [radius = 0.1];
        \fill (1.0, 0.0) circle [radius = 0.1];
        \fill (0.5, 0.866) circle [radius = 0.1];
        \fill (1.5, 0.866) circle [radius = 0.1];
        \fill (1.0, 1.732) circle [radius = 0.1];
        \fill (0.0, 0.0) node [below] {$(0, 0)$};
        \fill (0.5, 0.866) node [left] {$(1, 1)$};
        \fill (1.0, 0.0) node [below] {$(1, 0)$};
        \draw (0.0, 0.0) -- (0.5, 0.866);
        \draw (0.5, 0.866) -- (1.5, 0.866);
        \draw (1.0, 1.732) -- (1.5, 0.866);
        \draw [dashed] [red] (1.5, 0.866) -- (1.0, 0.0);
        \draw [dashed] [red] (1.0, 1.732) -- (0.5, 0.866);
        \draw [dashed] [red] (1.0, 1.732) -- (0.5, 0.866);
      \end{tikzpicture}

@@

    \end{tabular}
    \caption{Two conformations when a bead at (1,0) has a free hand.}
    \label{fig:beadat(1,0)}
  \end{center}
	\end{minipage}%	
\hspace{1cm}\begin{minipage}{.45\textwidth}
		\begin{center}
			\begin{tabular}{cc}
				\begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
				\fill (0.0, 0.0) circle [radius = 0.1];
				\fill (0.5, 0.866) circle [radius = 0.1];
				\fill (-0.5, 0.866) circle [radius = 0.1];
				\fill (0.0, 1.732) circle [radius = 0.1];
				\fill (0.0, 0.0) node [below] {$(1, 1)$};
				\draw (0.0, 0.0) -- (0.5, 0.866);
				\draw (0.5, 0.866) -- (-0.5, 0.866);
				\draw (0.0, 1.732) -- (-0.5, 0.866);
				\draw [dashed] [red] (0.0, 1.732) -- (0.5, 0.866);
				\draw [dashed] [red] (0.0, 0.0) -- (-0.5, 0.866);
				\end{tikzpicture}

@@

			\end{tabular}
			\caption{Two conformations when a bead $i$ at (1,1) has a free hand}
			\label{fig:beadi2}
		\end{center}
	\end{minipage}

\end{figure}

%
%\begin{figure}
%  \begin{center}
%    \begin{tabular}{cc}
%      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
%        \fill (0.0, 0.0) circle [radius = 0.1];
%        \fill (0.5, 0.866) circle [radius = 0.1];
%        \fill (-0.5, 0.866) circle [radius = 0.1];
%        \fill (0.0, 1.732) circle [radius = 0.1];
%        \fill (0.0, 0.0) node [below] {$(1, 1)$};
%        \draw (0.0, 0.0) -- (0.5, 0.866);
%        \draw (0.5, 0.866) -- (-0.5, 0.866);
%        \draw (0.0, 1.732) -- (-0.5, 0.866);
%        \draw [dashed] [red] (0.0, 1.732) -- (0.5, 0.866);
%        \draw [dashed] [red] (0.0, 0.0) -- (-0.5, 0.866);
%      \end{tikzpicture}
%
%      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
%        \fill (0.0, 0.0) circle [radius = 0.1];
%        \fill (-0.5, 0.866) circle [radius = 0.1];
%        \fill (0.5, 0.866) circle [radius = 0.1];
%        \fill (0.0, 1.732) circle [radius = 0.1];
%        \fill (0.0, 0.0) node [below] {$(1, 1)$};
%        \draw (0.0, 0.0) -- (-0.5, 0.866);
%        \draw (-0.5, 0.866) -- (0.5, 0.866);
%        \draw (0.0, 1.732) -- (0.5, 0.866);
%        \draw [dashed] [red] (0.0, 0.0) -- (0.5, 0.866);
%        \draw [dashed] [red] (0.0, 1.732) -- (-0.5, 0.866);
%      \end{tikzpicture}
%
%    \end{tabular}
%    \caption{Two conformations when a bead $i$ at (1,1) has a free hand}
%    \label{fig:beadi2}
%  \end{center}
%\end{figure}

@@

%\end{enumerate}
